---
title: "Lab 5"
output: html_document
author: Jayant Shah
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load data

```{r}
acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)

pers <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)
```

# 1.

```{r}
pers1 = subset(pers, pers$INJ_SEV==4)
```

# 2.

```{r}
library(dplyr)
pers1_MAKE = pers1 %>% group_by(STATE,MAKE) %>% summarise(vehicle_make_acc=n())
# Delete NA
pers1_MAKE = na.omit(pers1_MAKE)

pers1_MAKE1 = pers1_MAKE %>% group_by(STATE) %>% filter(vehicle_make_acc==max(vehicle_make_acc)) %>% as.data.frame
pers1_MAKE1
```

# 3.

```{r}
state = readxl::read_xlsx("FRPP_GLC_-_United_StatesNov42021.xlsx")
names(state)[3] = "State_Code"
df = merge(pers1_MAKE1, state, by.x = "STATE", by.y = "State_Code", all.x = T, all.y = T)
df = na.omit(df[,c(1,2,3,5)])
df = df[!duplicated(df[,3:4]), ]


library(usmap)
library(tidyverse)

us <- usmap::us_map()
region = df$`State Name`

region <- str_to_title(region)
share =  df$MAKE
share_df <- data.frame(region, share)

us_val <- 
  left_join(us, share_df, by = c("full" ="region")) 
#> Warning: Column `full`/`region` joining character vector and factor, coercing
#> into character vector

us_centroids <- 
  us_val %>%
  group_by(full) %>% 
  summarise(centroid.x = mean(range(x)), 
            centroid.y = mean(range(y)),
            label = unique(toupper(str_sub(full,1,2))),
            share = unique(share))

ggplot() + 
  geom_polygon(data = us_val, 
               aes(x,y, group = group, fill = as.factor(share)), 
               color = "black",
               size = .1) +
  geom_text(data = us_centroids, 
            aes(centroid.x, centroid.y, label = paste(label, "\n", share)),
            size = 5/14*8) +
  scale_fill_brewer(name = "MAKE", 
                    palette = "Blues") +
  theme_void()
```

Most fatal accidents are associated with Chevrolet and Ford cars.

# 4.

```{r}
pers_acc = merge(pers, acc[,-c(1,4,10,12,13,16,17,19,20,29,30,40)], by.x = "ST_CASE", by.y ="ST_CASE")

```

# 5.

```{r}
df5_1 = pers_acc %>% group_by(DAY_WEEK) %>% summarise(acc_DAY_WEEK=n())

barplot( height=
           df5_1$acc_DAY_WEEK, names=df5_1$DAY_WEEK , density=c(5,10,20,30,7,15,11) , angle=c(0,45,90,11,36,50,80) , col="brown", main = "Number of accidents by day of the week", ylab = "Number of accidents", xlab = "Day of the week")
```

The greatest number of accidents occurs on Sunday and Saturday. The least number of accidents occurs on Thursday.

```{r}
df5_2 = pers_acc %>% group_by(HOUR) %>% summarise(acc_HOUR=n())
df5_2 = df5_2[1:24,]
barplot( height=
           df5_2$acc_HOUR, names=df5_2$HOUR ,border="#69b3a2", col="white", main = "Number of accidents by hour", ylab = "Number of accidents", xlab = "Hour")
```

The greatest number of accidents occurs from 17 to 19 hours. The smallest number of accidents occurs at 4 a.m.

```{r}
df5_3 = pers_acc %>% group_by(SEX) %>% summarise(acc_SEX=n())
df5_3 = df5_3[1:2,]
barplot( height=
           df5_3$acc_SEX, names=df5_3$SEX ,border="red", col="blue", density=10, main = "Number of accidents by sex", ylab = "Number of accidents", xlab = "Sex")
```

Most accidents happen to men.

# 6.

```{r}
pers1_county = pers1 %>% group_by(COUNTY) %>% summarise(value=n())

names(state)[6]="County"
names(state)[11]="Station_Code"

df2 = merge(pers1_county, state[,c(6,11)], by.x = "COUNTY", by.y = "County", all.x = F, all.y = F)


df2$Code = substr(df2$Station_Code,1,5)



library(socviz)
library(ggthemes)

county_full <- left_join(county_map, county_data, by = "id")

df3 = merge(county_full, df2[,c(2,4)], by.x = "id", by.y = "Code", all.x = F, all.y = F)

df3$value_dens = ifelse(df3$value<10,"0-10",ifelse(df3$value<50,"10-50",ifelse(df3$value<100,"50-100",ifelse(df3$value<500,"100-500",">500"))))

p <- ggplot(data = df3,
            mapping = aes(x = long, y = lat,
                          fill = value_dens, 
                          group = group))

p1 <- p + geom_polygon(color = "gray90", size = 0.05) + coord_equal()

p2 <- p1 + scale_fill_brewer(palette="Blues",
                             labels = c("0-10", "10-50", "50-100", "100-500",
                                         ">500"))

p2 + labs(fill = "Number of deaths") +
    theme_map() +
    guides(fill = guide_legend(nrow = 1)) + 
    theme(legend.position = "bottom")




```

# 7.

```{r}
pers1$Sum.Win = ifelse(pers1$MONTH==12|pers1$MONTH==1|pers1$MONTH==2, "Winter", ifelse(pers1$MONTH==6|pers1$MONTH==7|pers1$MONTH==8, "Summer",NA))

pers1_SW = pers1 %>% group_by(STATE, Sum.Win) %>% summarise(value=n())
pers1_SW = na.omit(pers1_SW)

df4 = merge(pers1_SW, state, by.x = "STATE", by.y = "State_Code", all.x = T, all.y = T)
df4 = na.omit(df4[,c(1,2,3,5)])
df4 = df4[!duplicated(df4[,3:4]), ]

ggplot(data=df4, aes(x=`State Name`, y=value, fill=Sum.Win)) +
geom_bar(stat="identity", color="black", position=position_dodge())+coord_flip()+
  theme(text = element_text(size=7))+ scale_fill_discrete(name = "Season")

```

As you can see, the number of fatal accidents depends on the season and state.








